# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r loaddata}
setwd("c:/jay/rstuff/RepData_PeerAssessment1")
library(lattice)
library(ggplot2)

aData <- read.csv("activity.csv")
aData$date <- as.Date(aData$date)
```


## What is mean total number of steps taken per day?
```{r totalsteps}
tSteps <- aggregate(steps ~ date, data = aData, sum)
hist(tSteps$steps,xlab="Steps",main="Histogram of Total Number of Steps Taken Each Day")

mn <- sprintf("%5.2f",mean(tSteps$steps))
mdn <- median(tSteps$steps)
```
1. The mean total number of steps taken per day is `r mn`
1. The median total number of steps taken per day is `r mdn`  

## What is the average daily activity pattern?
```{r avgdaily}
int_tSteps <- aggregate(steps ~ interval, data = aData, mean)

#Make Timestamps Prettier
ts = seq(1, 288, 48)

int_tSteps$timestamp <- strptime(sprintf("%04d",int_tSteps$interval),"%H%M")
scalesList <- list(x = list(labels = format(int_tSteps$timestamp[ts],"%H:%M"), at = int_tSteps$interval[ts]))

xyplot(steps ~ interval, data = int_tSteps, type ="l", main="Average Number Of Steps Taken", scales = scalesList)


maxsteps <- format(int_tSteps[which.max(int_tSteps$steps),3],"%H:%M")
maxsteps2 <- format(int_tSteps[which.max(int_tSteps$steps) + 1,3],"%H:%M")
maxstepsd <- sprintf("%3.2f", int_tSteps[which.max(int_tSteps$steps),2])
```
1. The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is `r maxsteps` to `r maxsteps2` with `r maxstepsd` steps

## Imputing missing values
```{r missvalues}
mv <- sum(is.na(aData$steps))

# Let's Use Mean Steps For Each 5-minute interval
meanSteps <- aggregate(steps ~ interval, data = aData, mean)
names(meanSteps)[2] <- "asteps"
adj_aData <- merge(aData,meanSteps,by = "interval")
adj_aData$steps <- ifelse(is.na(adj_aData$steps),adj_aData$asteps,adj_aData$steps)

adj_tSteps <- aggregate(steps ~ date, data = adj_aData, sum)
amn <- sprintf("%5.2f", mean(adj_tSteps$steps))
amdn <- sprintf("%5.2f", median(adj_tSteps$steps))

hist(adj_tSteps$steps,xlab="Steps",main="Histogram of Adjusted Total Number of Steps Taken Each Day")
```

**The missing values have been replace by the total mean for corresponding interval**

1. The  total number of missing values in the dataset is `r mv`
1. The adjusted mean total number of steps taken per day is `r amn`.
1. The adjusted median total number of steps taken per day is `r amdn`.
1. As compared to the previous dataset, the mean has not changed, and the median has minimally increased.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekend}
adj_aData$whatday <- ifelse(weekdays(adj_aData$date) %in% c("Saturday", "Sunday"),"weekend","weekday")
adj_int_tSteps <- aggregate(steps ~ interval + whatday, data = adj_aData, mean)

#Make Intervals Prettier
adj_int_tSteps$timestamp <- strptime(sprintf("%04d",adj_int_tSteps$interval),"%H%M")
scalesList <- list(x = list(labels = format(adj_int_tSteps$timestamp[ts],"%H:%M"), at = adj_int_tSteps$interval[ts]))


xyplot(steps ~ interval | whatday, data = adj_int_tSteps, layout = c(1,2), type ="l", main="Average Number Of Steps Taken", scales = scalesList)
```